Precise predictions for same-sign W-bosons scattering at the LHC

### 2018-02-20 Draft/scanNLO.tex
Many changes from Ansgar and Mathieu

These differences are much more evident in differential distributions.
In Fig.~\ref{fig:mjjdyjj_1d_1}, we show the distributions in the di-jet invariant mass $m_{\Pj\Pj}$ and rapidity separation $|\Delta y_{\Pj\Pj}|$.
For large $m_{\Pj\Pj}$ and large $|\Delta y_{\Pj\Pj}|$, as expected, the VBS approximation is performing well and its $s$-channel extension agree with the full calculation within $10\%$ per cent.
% On the other hand, for low invariant and low rapidity separation the the VBS approximation is failing at a level of at least $30\%$.
% In this very same region, the VBS approximation with $s$-channel is good within $10\%$.
This is in contrast with the region  $200 \GeV < m_{\Pj\Pj} < 500 \GeV$ and $2<|\Delta y_{\Pj\Pj}|<2.5$, the discrepancy between the $|t|^2 + |u|^2$ approximation and the full computation goes up to $30\%$.
The inclusion of $s$-channels cures partly the discrepancy in this region.
Still, for the very low $m_{\Pj\Pj}$ a difference of about $5\%$ remains.
This might indicate that also interferences (GP: and non--factorizable QCD corrections ? \MP{I don't know, we have to think about it}) are needed in this phase-space region.

@@

\begin{figure}[hbt]
\centering
{\includegraphics[scale=0.35]{figures/scanfigures/mjj_nlo.pdf}}
{\includegraphics[scale=0.35]{figures/scanfigures/dyjj_nlo.pdf}}
\caption{
Differential distributions in $m_{\Pj\Pj}$ (left) and $|\Delta y_{\Pj\Pj}|$ (right) at NLO QCD \emph{i.e.}\ at order $\mathcal{O}(\alphas\alpha^6)$ for the full computation and two approximations.
In addition to the cuts of Sec.~\ref{subsec:inputpar}, the VBS cuts take the values: $m_{\Pj\Pj}>200 \GeV$ and $|\Delta y_{\Pj\Pj}|>2$.} 
\label{fig:mjjdyjj_1d_1}
\end{figure}

\begin{figure}[hbt]
\centering
{\includegraphics[scale=0.39]{figures/scanfigures/a6as_vbfnloVSrecola_tu.pdf}}
\caption{Ratio of cross sections (fb) per bin of $(m_{\Pj\Pj},\,|\Delta y_{\Pj\Pj}|)$ at NLO QCD \emph{i.e.}\ at order $\mathcal{O}(\alphas\alpha^6)$ for the VBS approximation over the full computation.
In addition to the cuts of Sec.~\ref{subsec:inputpar}, the VBS cuts take the values: $m_{\Pj\Pj}>200 \GeV$ and $|\Delta y_{\Pj\Pj}|>2$.}
\label{fig:ratio2d_NLO}
\end{figure}
%
\begin{figure}[hbt]
\centering
{\includegraphics[scale=0.39]{figures/scanfigures/a6as_vbfnloVSrecola_stu.pdf}}
\caption{Ratio of cross sections (fb) per bin of $(m_{\Pj\Pj},\,|\Delta y_{\Pj\Pj}|)$ at NLO QCD \emph{i.e.}\ at order $\mathcal{O}(\alphas\alpha^6)$ for the VBS approximation with $s$-channel contributions over the full computation.
In addition to the cuts of Sec.~\ref{subsec:inputpar}, the VBS cuts take the values: $m_{\Pj\Pj}>200 \GeV$ and $|\Delta y_{\Pj\Pj}|>2$.}
\label{fig:mjjdyjj_2d_NLO}
\end{figure}

As expected, in the low invariant mass--low rapidity separation region of the jet pair the VBS approximation fails significantly (up to $40\%$ discrepancies).
The inclusion of the $s$-channel brings the difference down to at most $5\%$.
However, the positive discrepancy shown in the low $m_{\Pj\Pj}$ region (black curve on the upper plots of Fig.~\ref{fig:mjjdyjj_1d_1}) can be traced back to the low $m_{\Pj\Pj}$, large $\Delta y_{\Pj\Pj}$ region of Fig.~\ref{fig:mjjdyjj_2d_NLO}.
In this region, the two leading jets have soft transverse momenta, according to the following low-angle approximation

\begin{equation}
\begin{split}
m_{\Pj_i\Pj_j}^2 &= 2\,\ptsub{\Pj_i}\ptsub{\Pj_j}\,(\cosh \Delta y_{\Pj_i\Pj_j} - \cos \Delta\phi_{\Pj_i\Pj_j}) \\
&\approx 2\,\ptsub{\Pj_i}\ptsub{\Pj_j}\,\cosh \Delta y_{\Pj_i\Pj_j} \,\,.
\end{split}
\end{equation}

The same positive discrepancy for the $|s|^2 + |t|^2 + |u|^2$ approximation, can be seen in the low $\pt$ region of the leading jet in the upper plot of Fig.~\ref{fig:mjjdyjj_1d_2}.
In the large invariant mass--small rapidity separation region of Fig.~\ref{fig:mjjdyjj_2d_NLO}, discrepancies at the level of $15\%$ are present.
This can be traced back to the large $\pt$ and central rapidity region of the leading jets kinematics, shown in Fig.~\ref{fig:mjjdyjj_1d_2}.
For such distributions, despite the $s$-channel inclusion, the discrepancy between the approximated and full result is of about $5-10\%$.
In the VBS signal-region the VBS approximation shows a good agreement with the full calculation as documented in details below.

\begin{figure}[hbt]
\centering
{\includegraphics[scale=0.35]{figures/scanfigures/ptj1_nlo.pdf}}
{\includegraphics[scale=0.35]{figures/scanfigures/yj1_nlo.pdf}}
\caption{Differential distributions in the transverse momentum and rapidity of the hardest tagging jet at NLO QCD \emph{i.e.}\ at order $\mathcal{O}(\alphas\alpha^6)$ for the full computation and two approximations.
In addition to the cuts of Sec.~\ref{subsec:inputpar}, the VBS cuts take the values: $m_{\Pj\Pj}>200 \GeV$ and $|\Delta y_{\Pj\Pj}|>2$.} 
\label{fig:mjjdyjj_1d_2}
\end{figure}

Concerning leptonic observables, we show in Fig.~\ref{fig:mjjdyjj_1d_3} the distributions of the lepton-lepton invariant mass and of the Zeppenfeld variable of the electron. 
It is defined as
%
\begin{equation}
%  z_{\Pe^+} = \frac{y_{\Pe^+}-\frac{y_{\Pj_1}+y_{\Pj2}}2}{|y_{\Pj_1}-y_{\Pj_2}|} \,.
  z_{\Pe^+} = \frac{y_{\Pe^+}-\frac{y_{\Pj_1}+y_{\Pj2}}2}{|\Delta y_{jj}|} \,.
  \label{eq:Zeppenfeld}
\end{equation}
%
Analogous definitions will later also be used for the Zeppenfeld variable of the muon and of the third jet.
%
\begin{figure}[hbt]
\centering
{\includegraphics[scale=0.35]{figures/scanfigures/mll_nlo.pdf}}
{\includegraphics[scale=0.35]{figures/scanfigures/zel_nlo.pdf}}
\caption{Differential distributions in the lepton-lepton invariant mass and the electron Zeppenfeld variable at NLO QCD \emph{i.e.}\ at order $\mathcal{O}(\alphas\alpha^6)$ for the full computation and two approximations.
In addition to the cuts of Sec.~\ref{subsec:inputpar}, the VBS cuts take the values: $m_{\Pj\Pj}>200 \GeV$ and $|\Delta y_{\Pj\Pj}|>2$.} 
\label{fig:mjjdyjj_1d_3}
\end{figure}
%
The {\sc VBFNLO} result for the $e^+\mu^+$ invariant mass agrees rather well with the full curve, obtained from {\sc MoCa\-NLO+Recola}.
The prediction from {\sc Bonsay} is about $10\%$ lower.
The discrepancies are roughly constant over the whole spectrum.
Instead, the right panel of Fig.~\ref{fig:mjjdyjj_1d_3} clearly shows that the Zeppenfeld variable of the positron $z_e$ is strongly affected by the exclusion of $s$-channels, with increasing discrepancy with respect to the full result at large values.
The muon observable $z_{\mu}$ behaves identically to the electron one, $z_e$.

>>>>

These differences are much more evident in differential distributions.
In Fig.~\ref{fig:mjjdyjj_1d_1}, we show the distributions in the di-jet invariant mass $m_{\Pj\Pj}$ and rapidity separation $|\Delta y_{\Pj\Pj}|$.
For large $m_{\Pj\Pj}$ and large $|\Delta y_{\Pj\Pj}|$, as expected, the VBS approximation is performing well and its $s$-channel extension agree with the full calculation within $10\%$ per cent.
% On the other hand, for low invariant and low rapidity separation the the VBS approximation is failing at a level of at least $30\%$.
% In this very same region, the VBS approximation with $s$-channel is good within $10\%$.
This is in contrast with the region  $200 \GeV < m_{\Pj\Pj} < 500 \GeV$ and $2<|\Delta y_{\Pj\Pj}|<2.5$, the discrepancy between the $|t|^2 + |u|^2$ approximation and the full computation goes up to $30\%$.
The inclusion of $s$-channels cures partly the discrepancy in this region.
Still, for the very low $m_{\Pj\Pj}$ a difference of about $5\%$ remains \MP{To be checked the exact numerical value in data}.
This might indicate that interference contributions and/or non-factorisable QCD corrections play a non-negligible role in this phase-space region.

@@

\begin{figure*}[hbt]
\centering
{\includegraphics[scale=0.35]{figures/scanfigures/mjj_nlo.pdf}}
{\includegraphics[scale=0.35]{figures/scanfigures/dyjj_nlo.pdf}}
\caption{
Differential distributions in the di-jet invariant mass (left) and the rapidity-separation of the two tagging jets (right) at NLO QCD \emph{i.e.}\ at order $\mathcal{O}(\alphas\alpha^6)$ for the full computation and two approximations.
The upper plots provide the absolute value for each prediction while the lower plots presents all predictions normalised to {\sc MoCaNLO}+{\sc Recola} which is one of the full predictions.
In addition to the cuts of Sec.~\ref{subsec:inputpar}, the VBS cuts take the values: $m_{\Pj\Pj}>200 \GeV$ and $|\Delta y_{\Pj\Pj}|>2$.} 
\label{fig:mjjdyjj_1d_1}
\end{figure*}

\begin{figure*}[hbt]
\centering
{\includegraphics[scale=0.395]{figures/scanfigures/a6as_vbfnloVSrecola_tu.pdf}}
{\includegraphics[scale=0.395]{figures/scanfigures/a6as_vbfnloVSrecola_stu.pdf}}
\caption{Ratio of cross sections per bin in the plan $\left(m_{\Pj\Pj}, |\Delta y_{\Pj\Pj}| \right)$ at NLO QCD \emph{i.e.}\ at order $\mathcal{O}(\alphas\alpha^6)$ for the VBS approximation over the full computation.
Ratio of approximated squared amplitudes over the full matrix element.
The approximated squared amplitudes are computed as $|\mathcal{A}|^2 \sim |t|^2 + |u|^2$ (left) and $|\mathcal{A}|^2 \sim |s|^2 + |t|^2 + |u|^2$ (right)
In addition to the cuts of Sec.~\ref{subsec:inputpar}, the VBS cuts take the values: $m_{\Pj\Pj}>200 \GeV$ and $|\Delta y_{\Pj\Pj}|>2$.}
\label{fig:ratio2d_NLO}
\end{figure*}
%
% \begin{figure}[hbt]
% \centering
% {\includegraphics[scale=0.39]{figures/scanfigures/a6as_vbfnloVSrecola_stu.pdf}}
% \caption{Ratio of cross sections (fb) per bin in the plan $\left(m_{\Pj\Pj}, |\Delta y_{\Pj\Pj}| \right)$ at NLO QCD \emph{i.e.}\ at order $\mathcal{O}(\alphas\alpha^6)$ for the VBS approximation with $s$-channel contributions over the full computation.
% In addition to the cuts of Sec.~\ref{subsec:inputpar}, the VBS cuts take the values: $m_{\Pj\Pj}>200 \GeV$ and $|\Delta y_{\Pj\Pj}|>2$.}
% \label{fig:mjjdyjj_2d_NLO}
% \end{figure}

As expected, in the low invariant mass--low rapidity separation region of the jet pair the VBS approximation fails significantly (up to $40\%$ discrepancies).
The inclusion of the $s$-channel brings the difference down to at most $5\%$ \MP{To be checked the exact numerical value in data}.
However, the positive discrepancy shown in the low $m_{\Pj\Pj}$ region (black curve on the upper plots of Fig.~\ref{fig:mjjdyjj_1d_1}) can be traced back to the low $m_{\Pj\Pj}$, large $\Delta y_{\Pj\Pj}$ region of Fig.~\ref{fig:mjjdyjj_2d_NLO}.
% In this region, the two leading jets have soft transverse momenta, according to the following low-angle approximation
% 
% \begin{equation}
% \begin{split}
% m_{\Pj_i\Pj_j}^2 &= 2\,\ptsub{\Pj_i}\ptsub{\Pj_j}\,(\cosh \Delta y_{\Pj_i\Pj_j} - \cos \Delta\phi_{\Pj_i\Pj_j}) \\
% &\approx 2\,\ptsub{\Pj_i}\ptsub{\Pj_j}\,\cosh \Delta y_{\Pj_i\Pj_j} \,\,.
% \end{split}
% \end{equation}

The same positive discrepancy for the $|s|^2 + |t|^2 + |u|^2$ approximation, can be seen in the low $\pt$ region of the leading jet in the upper plot of Fig.~\ref{fig:mjjdyjj_1d_2}.
In the large invariant mass--small rapidity separation region of Fig.~\ref{fig:mjjdyjj_2d_NLO}, discrepancies at the level of $15\%$ are present.
This can be traced back to the large $\pt$ and central rapidity region of the leading jets kinematics, shown in Fig.~\ref{fig:mjjdyjj_1d_2}.
For such distributions, despite the $s$-channel inclusion, the discrepancy between the approximated and full result is about $5-10\%$.
In the VBS signal-region the VBS approximation shows a good agreement with the full calculation as documented in details below.

\begin{figure*}[hbt]
\centering
{\includegraphics[scale=0.35]{figures/scanfigures/ptj1_nlo.pdf}}
{\includegraphics[scale=0.35]{figures/scanfigures/yj1_nlo.pdf}}
\caption{Differential distributions in the transverse momentum (left) and rapidity of the hardest tagging jet (right) at NLO QCD \emph{i.e.}\ at order $\mathcal{O}(\alphas\alpha^6)$ for the full computation and two approximations.
The upper plots provide the absolute value for each prediction while the lower plots presents all predictions normalised to {\sc MoCaNLO}+{\sc Recola} which is one of the full predictions.
In addition to the cuts of Sec.~\ref{subsec:inputpar}, the VBS cuts take the values: $m_{\Pj\Pj}>200 \GeV$ and $|\Delta y_{\Pj\Pj}|>2$.} 
\label{fig:mjjdyjj_1d_2}
\end{figure*}

Concerning leptonic observables, we show in Fig.~\ref{fig:mjjdyjj_1d_3} the distributions of the lepton-lepton invariant mass and of the Zeppenfeld variable of the electron, defined as
%
\begin{equation}
%  z_{\Pe^+} = \frac{y_{\Pe^+}-\frac{y_{\Pj_1}+y_{\Pj2}}2}{|y_{\Pj_1}-y_{\Pj_2}|} \,.
  z_{\Pe^+} = \frac{y_{\Pe^+}-\frac{y_{\Pj_1}+y_{\Pj2}}2}{|\Delta y_{jj}|} \,.
  \label{eq:Zeppenfeld}
\end{equation}
%
Analogous definitions will later also be used for the Zeppenfeld variable of the muon and of the third jet.
%
\begin{figure*}[hbt]
\centering
{\includegraphics[scale=0.35]{figures/scanfigures/mll_nlo.pdf}}
{\includegraphics[scale=0.35]{figures/scanfigures/zel_nlo.pdf}}
\caption{Differential distributions in the lepton-lepton invariant mass (left) and the electron Zeppenfeld variable (right) at NLO QCD \emph{i.e.}\ at order $\mathcal{O}(\alphas\alpha^6)$ for the full computation and two approximations.
The upper plots provide the absolute value for each prediction while the lower plots presents all predictions normalised to {\sc MoCaNLO}+{\sc Recola} which is one of the full predictions.
In addition to the cuts of Sec.~\ref{subsec:inputpar}, the VBS cuts take the values: $m_{\Pj\Pj}>200 \GeV$ and $|\Delta y_{\Pj\Pj}|>2$.} 
\label{fig:mjjdyjj_1d_3}
\end{figure*}
%
The {\sc VBFNLO} result for the $e^+\mu^+$ invariant mass agrees rather well with the full curve, obtained from {\sc MoCa\-NLO+Recola}.
The prediction from {\sc Bonsay} is about $10\%$ lower.
The discrepancies are roughly constant over the whole spectrum.
Instead, the right panel of Fig.~\ref{fig:mjjdyjj_1d_3} clearly shows that the Zeppenfeld variable of the positron $z_e$ is strongly affected by the exclusion of $s$-channels, with increasing discrepancy with respect to the full result at large values.
The muon observable $z_{\mu}$ behaves identically to the electron one, $z_e$.

### 2018-02-20 Draft/shower.tex
Many changes from Ansgar and Mathieu

We now discuss how different predictions compare when the matching to parton-shower (PS) is included. For such
a comparison we expect larger discrepancy than what we found at fixed-order, as a consequence of the different
matching schemes, parton shower employed and of other details of the matching (such as the choice of the parton shower initial scale). Among
the codes capable of providing fixed-order results, presented before, {\sc MG5\_aMC}, the {\sc Powheg-Box}, and {\sc VBFNLO}
can also provide results at (N)LO+PS accuracy. For {\sc VBFNLO}, we
restrict ourselves to showing results only in the VBS approximation,
{\emph i.e.}\ the $s$-channel contributions are neglected here. Besides,
also {\sc PHANTOM} is employed for LO+PS results.\\
{\sc MG5\_aMC}, which
employs the {\sc MC@NLO}~\cite{Frixione:2002ik} matching procedure, will be used together with {\sc Pythia8}~\cite{Sjostrand:2014zea} (version 8.223)
and {\sc Herwig++}~\cite{Bahr:2008pv, Bellm:2013hwb} (version 2.7.1). For the {\sc Powheg-Box}, the eponymous 
\MP{Is it the correct word? Because it would mean that the word ``Powheg'' has different meaning. Isn't it?} 
\AK{Changed to epononymous. Although still not correct as the powheg-box takes its name from the powheg matching and not the other way around... :)}
matching procedure is employed~\cite{Nason:2004rx,Frixione:2007vw}, together with {\sc Pythia8} (version 8.230). {\sc VBFNLO} serves as a matrix-element and phase-space provider
for the {\sc Matchbox} module~\cite{Platzer:2011bc} of {\sc
Herwig7}~\cite{Bellm:2015jjp,Bellm:2017bvx}, using an extended version of the BLHA
interface~\cite{Binoth:2010xt,Alioli:2013nda,Andersen:2014efa}. The {\sc Matchbox} module makes it
possible to choose between {\sc MC\-@NLO}-like and {\sc Powheg}-like
matching. As parton shower, both the default angular-ordered shower as
well as the dipole shower can be employed.
Finally, {\sc PHANTOM} results will be shown matched with {\sc Pythia 8} and {\sc Herwig++}.\AK{versions??}
Whenever {\sc Pythia8}\ is used, the Monash tune~\cite{Skands:2014pea} is selected.

Results will be presented within the cuts described in Sec.~\ref{subsec:inputpar}, applied after shower and hadronisation (this implies that jets
are obtained by clustering stable hadrons, and not QCD partons). It follows that at the event-generation level, looser cuts (or even no cuts at all)
must be employed in order not to bias the results. {\bf MZ lepton-jet separation at the hard-event level?} \AK{In powheg the separation is done after shower. So it's really the jet.}.

Compared to the fixed order computations, a slightly different set-up has been employed for {\sc MG5\_aMC} in order to simplify the calculation: instead of generating the full
${\rm p}{\rm p}\to\mu^+\nu_\mu{\rm e}^+\nu_{\rm e}{\rm j}{\rm j}$ process, since it is anyway dominated by doubly-resonant contribution, the
events are produced for the process with two stable W$^+$ bosons (${\rm p}{\rm p}\to{\rm W^+}{\rm W^+}{\rm j}{\rm j}$), and these W$^+$ bosons
are decayed with {\sc MadSpin}~\cite{Artoisenet:2012st} (keeping spin correlations) before the PS. Since {\sc MadSpin}\ computes
the partial and total decay width of the W bosons at LO accuracy only, while in Section~\ref{subsec:inputpar} the NLO width is employed,
a small effect (6\%) on the normalisation of distribution is induced. Finally, when the renormalisation
and factorisation scales are set, the $\Delta R_{\Pj\Pl}$ cut is not imposed during the jet-clustering procedure, but this has no visible effect
on the results \MP{Is it true? Because I remember some numbers from Christopher with $1\%$ or so}.

>>>>

We now discuss how different predictions compare when the matching to parton-shower (PS) is included. For such
a comparison we expect larger discrepancy than what we found at fixed-order, as a consequence of the different
matching schemes, parton shower employed and of other details of the matching (such as the choice of the parton shower initial scale). Among
the codes capable of providing fixed-order results, presented before, {\sc MG5\_aMC}, the {\sc Powheg-Box}, and {\sc VBFNLO}
can also provide results at (N)LO+PS accuracy. For {\sc VBFNLO}, we
restrict ourselves to showing results only in the VBS approximation,
{\emph i.e.}\ the $s$-channel contributions are neglected here. Besides,
also {\sc PHANTOM} is employed for LO+PS results.\\
{\sc MG5\_aMC}, which
employs the {\sc MC@NLO}~\cite{Frixione:2002ik} matching procedure, will be used together with {\sc Pythia8}~\cite{Sjostrand:2014zea} (version 8.223)
and {\sc Herwig++}~\cite{Bahr:2008pv, Bellm:2013hwb} (version 2.7.1). For the {\sc Powheg-Box}, the namesake
matching procedure is employed~\cite{Nason:2004rx,Frixione:2007vw}, together with {\sc Pythia8} (version 8.230). {\sc VBFNLO} serves as a matrix-element and phase-space provider
for the {\sc Matchbox} module~\cite{Platzer:2011bc} of {\sc
Herwig7}~\cite{Bellm:2015jjp,Bellm:2017bvx}, using an extended version of the BLHA
interface~\cite{Binoth:2010xt,Alioli:2013nda,Andersen:2014efa}. The {\sc Matchbox} module makes it
possible to choose between {\sc MC\-@NLO}-like and {\sc Powheg}-like
matching. As parton shower, both the default angular-ordered shower as
well as the dipole shower can be employed.
Finally, {\sc PHANTOM} results will be shown matched with {\sc Pythia 8} and {\sc Herwig++}.\AK{versions??}
Whenever {\sc Pythia8}\ is used, the Monash tune~\cite{Skands:2014pea} is selected.

Results will be presented within the cuts described in Sec.~\ref{subsec:inputpar}, applied after shower and hadronisation (this implies that jets
are obtained by clustering stable hadrons, and not QCD partons). It follows that at the event-generation level, looser cuts (or even no cuts at all)
must be employed in order not to bias the results. \ZM{lepton-jet separation at the hard-event level?} \AK{In powheg the separation is done after shower. So it's really the jet.}.

Compared to the fixed order computations, a slightly different set-up has been employed for {\sc MG5\_aMC} in order to simplify the calculation: instead of generating the full
${\rm p}{\rm p}\to\mu^+\nu_\mu{\rm e}^+\nu_{\rm e}{\rm j}{\rm j}$ process, since it is anyway dominated by doubly-resonant contribution, the
events are produced for the process with two stable W$^+$ bosons (${\rm p}{\rm p}\to{\rm W^+}{\rm W^+}{\rm j}{\rm j}$), and these W$^+$ bosons
are decayed with {\sc MadSpin}~\cite{Artoisenet:2012st} (keeping spin correlations) before the PS. Since {\sc MadSpin}\ computes
the partial and total decay width of the W bosons at LO accuracy only, while in Section~\ref{subsec:inputpar} the NLO width is employed,
a small effect (6\%) on the normalisation of distribution is induced. Finally, when the renormalisation
and factorisation scales are set, the $\Delta R_{\Pj\Pl}$ cut is not imposed during the jet-clustering procedure.
But this effect is below a per cent at the level of the fiducial cross sections.
